Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-17 MDPI_submission.Rmd
Commit 2023-12-17 21:10

# Introduction.

@@

We calculated the linear trends of SDR, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
We observe a positive trend for all-sky conditions
($`r alltrend`\,\%/y$), a very close but smaller trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloudy-skies ($`r cloudtrend`\,\%/y$).
In the studied period, there is no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR slope, around 1980 [@Wild2021; @Yuan2021; @Ohmura2009], a few years before the start of our records.
It is interesting to note, that for the observations period, the trend of the TSI is
$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$,
and thus we can not attribute any major effect on SDR trend to Solar variability.

@@

## Effects of the solar zenith angle on SDR.

@@

For all-sky conditions the brightening effect of SDR (positive trend) is
stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].

>>>>

# Introduction

@@

We calculated the linear trends of SDR, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
In the studied period, there is no
significant break or change in the variability pattern of the time
series.
We observe a positive trend for all-sky conditions
($`r alltrend`\,\%/y$), a very close but smaller trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloudy-skies ($`r cloudtrend`\,\%/y$).
In the studied period, there is no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR slope, around 1980 [@Wild2021; @Yuan2021; @Ohmura2009], a few years before the start of our records.
It is interesting to note, that for the observations period, the trend of the TSI is
$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$,
and thus we can not attribute any major effect on SDR trend to Solar variability.

@@

## Effects of the solar zenith angle on SDR

@@

For all-sky conditions the brightening effect of SDR (positive trend) is
stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2})
ranging from about $0.1\,\%/y$ to
about $0.7\,\%/y$.
The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].

### 2023-12-17 MDPI_submission.tex
Commit 2023-12-17 21:10

\hypertarget{introduction.}{%
\section{Introduction.}\label{introduction.}}

@@

We calculated the linear trends of SDR, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\hypertarget{effects-of-the-solar-zenith-angle-on-sdr.}{%
\subsection{Effects of the solar zenith angle on
SDR.}\label{effects-of-the-solar-zenith-angle-on-sdr.}}

@@

For all-sky conditions the brightening effect of SDR (positive trend) is
stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
\ref{fig:szatrends-2}). The trends in the morning and afternoon hours
are more or less consistent with small differences, which can be
attributed to systematic diurnal variations of aerosols, particularly
during the warm period of the year \citep{Wang2021}.

>>>>

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

We calculated the linear trends of SDR, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In the studied
period, there is no significant break or change in the variability
pattern of the time series. We observe a positive trend for all-sky
conditions (\(0.38\,\%/y\)), a very close but smaller trend for
clear-skies (\(0.097\,\%/y\)) and a negative weaker trend for
cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is no
significant break or change in the variability of the time series. Other
studies for the European region reported a change of the SDR slope,
around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years before
the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\hypertarget{effects-of-the-solar-zenith-angle-on-sdr}{%
\subsection{Effects of the solar zenith angle on
SDR}\label{effects-of-the-solar-zenith-angle-on-sdr}}

@@

For all-sky conditions the brightening effect of SDR (positive trend) is
stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
\ref{fig:szatrends-2}) ranging from about \(0.1\,\%/y\) to about
\(0.7\,\%/y\). The trends in the morning and afternoon hours are more or
less consistent with small differences, which can be attributed to
systematic diurnal variations of aerosols, particularly during the warm
period of the year \citep{Wang2021}.
